Block bootstrap consistency\\under weak assumptions

### 2016-06-16 Paper.tex
[#18] Make small text changes to reflect new assumption on means

\newpage
\noindent Block bootstraps, e.g.\ the Moving Blocks
(\citealp{Kun:89}, and \citealp{LiS:92}), Circular Block \citep{PoR:92}, and Stationary
Bootstraps \citep{PoR:94}, have become popular in Economics, partly
because they do not require the researcher to make parametric
assumptions about the data generating process.  They are valid under
general weak dependence and moment conditions.  Some recent papers
(\citealp{GoW:02}, and \citealp{GoJ:03}) relax the dependence and moment
conditions of the original papers to fit with the Near-Epoch-Dependence
(\ned) assumptions commonly used in econometrics.%
\footnote{\citet{GoW:02} show that these bootstrap methods can be
  applied to heterogeneous $L_{2+\delta}$-\ned\ processes of size
  $-2(r-1)/(r-2)$ on a strong mixing sequence of size
  $-r(2+\delta)/(r-2)$, where $r > 2$ and $\delta >0$, when the
  original series has uniformly bounded $3r$-moments.  \citet{GoJ:03}
  relax these conditions to $L_{2+\delta}$-\ned\ of size $-1$ and
  $r+\delta$ moments for the original series, and size
  $-(2+\delta)(r+\delta)/(r-2)$ for the underlying mixing series.
  Both papers require that the expected block length grow with $n$ and
  be $o(n^{1/2})$.  \cite{GoP:11} discuss these issues further.}%
\footnote{An array $\{X_{nt}\}$ is an $L_{\rho}$-\ned\ process on a
  mixing array $\{V_{nt}\}$ if
  \begin{equation}
    \| X_{nt} - \E(X_{nt}
    \mid V_{n,t-m},\dots,V_{n,t+m}) \|_{\rho} \leq d_{nt} v_m
  \end{equation}
  with $v_m \to 0$ as $m \to \infty$ and $\{d_{nt}\}$ an array of
  positive constants.  It is of size $-\gamma$ if $v_m = O(m^{-\gamma
    - \delta})$ for all $\delta>0$.  Dropping the index ``$n$'' gives
  the series definition.  Note that the underlying strong and uniform mixing arrays
  are not required to be stationary.} %
But these conditions are still stronger than
required for a \clt\ to hold; \citet{Jon:97} has established the \clt\
under $L_2$-\ned\ with smaller size and moment
restrictions.%
\footnote{\citet{Jon:97} proves that the \clt\ holds for averages of
  $L_2$-\ned\ processes of size $-1/2$ on a strong mixing series of
  size $-r/(r-2)$, $r > 2$ and the original series having bounded
  $r$-moments.} %
This paper shows that these block bootstrap
methods consistently estimate the distribution of the sample mean
under \citepos{Jon:97} assumptions and show that an \fclt\ holds as
well.%
\footnote{\citet{Rad:96} proves consistency for the Moving Blocks
  Bootstrap for any stationary strong mixing sequence that satisfies
  the \clt. This paper uses a similar method of proof to his, but also
  accommodates nonstationary sequences and the Stationary
  Bootstrap.} %
It also relaxes \citepos{GoW:02} and \citepos{GoJ:03} requirement that
the expected block length be $o(n^{1/2})$ to the original papers'
requirement that it be $o(n)$.

@@

\begin{thm}\label{T1}
  Suppose the following conditions hold.
  \begin{enumerate}
  \item $X_{nt}$ is $L_2$-\ned\ of size $-1/2$ on an array
    $\{V_{nt}\}$ that is either strong mixing of size $-r/(r-2)$ or
    uniform mixing of size $-r/2(r-1)$, with $r > 2$.  The
    \ned\ magnitude indices are denoted $\{d_{nt}\}$.
  \item The array $\mu_{nt} - \bar \mu_n$ is uniformly bounded
    and $\sum_{t=1}^n (\mu_{nt} - \bar \mu_n)^2 \to 0$,
    where $\E X_{nt} = \mu_{nt}$ and $\bar{\mu}_n = n^{-1} \sum_{t=1}^n
    \mu_{nt}$. Moreover, $\sqrt{n} \| \bar{X}_{n} - \bar\mu_n \|_2
    \to \sigma > 0$, with $\bar X_n = (1/n) \sum_{t=1}^n X_{nt}$.
  \item There exists an array of positive real numbers $\{c_{nt}\}$
    such that $(X_{nt} - \mu_{nt})/c_{nt}$ is uniformly $L_r$-bounded
    and $c_{nt}$ and $d_{nt}/c_{nt}$ are uniformly bounded in $n$
    and~$t$.
  \item $X_{nt}^{*}$ is generated by the Stationary Bootstrap with
    geometric block lengths with success probability $p_n$, $p_n = c
    n^{-a}$ and $a,c \in (0,1)$, or by the Moving or Circular Block
    bootstrap with block length $M_n$ such that $M_n \sim n^a$ for
    $a \in (0,1)$.  Let $M_{ni}$ be the block length of the $i$th
    block, $i=1,\dots,J_n$, and define $K_{n0} = 0$ and $K_{nj} =
    \sum_{i=1}^j M_{ni}$. The last block, $M_{n,J_n}$, is defined as
    $K_{n,J_n} - K_{n,J_n-1}$, so $K_{n,J_n} = n$ a.s.
  \end{enumerate}
  Then $\hat\sigma^{*2} \to^p \sigma^2$,
  \begin{equation}\label{eq:36}
    \sup_x \big\lvert \pr^{*}\big[\sqrt{n}(\bar X_{n}^{*} - \E^* \bar X_n^*) \leq x \big]
    - \pr \big[\sqrt{n}(\bar X_{n} - \E \bar X_n) \leq x \big] \big\rvert \to^p 0,
  \end{equation}
  and
  \begin{equation}
    \label{eq:2}
    \sup_x \big\lvert \pr^{*}\big[
    \sqrt{n}(\bar X_{n}^{*} -  \E^* \bar X_n^*) \big/ \hat\sigma_n^{*}
    \leq x \big] - \Phi(x) \big\rvert \to^p 0,
  \end{equation}
  where $\Phi$ is the \allcaps{cdf} of the Standard Normal
  distribution and
    \begin{equation}
      \label{eq:3}
      \hat{\sigma}_n^{*2} = \tfrac{1}{n} \sum_{j=1}^{J_n}
      \Big\{\sum_{t=K_{n,j-1}+1}^{K_{nj}} (X_{nt}^{*} - \bar X_{n}^{*})\Big\}^2.
    \end{equation}
\end{thm}

As mentioned earlier, these are the same size and mixing conditions
used by \citet{Jon:97}.
Note that De Jong does allow a little bit more flexibility in the
conditions on the array $\{c_{nt}\}$ \citep[see also][]{Dav:93};
essentially, he allows for a single set of blocks with the
maximal $\{c_{nt}\}$ over each block well-behaved, while this
paper requires this condition to hold for every possible partition of
blocks.  This additional restriction is required because the
Stationary Bootstrap will select the blocks randomly and is similar to
\citepos{JoD:00b} requirement for the \fclt. Similarly, our assumption
on the dispersion of the individual means, $(\mu_{nt} - \bar\mu_n)^2$, is
slightly stronger than \citepos{GoW:02} and \citepos{GoJ:03} to
accommodate larger block sizes.%
\footnote{Note that this is one reason that we are presenting our
  results using general triangular-array notation: to allow for the
  means to change over time for comparison with \citepos{GoW:02} and
  \citepos{GoJ:03} results.}

@@

Our insight is that the expectation of squared blocks of the bootstrap
process can be expressed as a sequence of \emph{contiguous} blocks of
the original process, so the arguments that establish convergence of
the original squared blocks can be applied with only minor changes to
the bootstrapped blocks. Consider the Moving Blocks Bootstrap,%
\footnote{To make this presentation as simple as possible, assume for
  now that $n = M_n J_n$ exactly.} %
for example, and let
\begin{equation*}
  Z_{nj}^* = \tfrac{1}{\sqrt{n}} \sum_{t=(j-1) M_n + 1}^{j M_n} (X_{nt}^* - \bar X_n).
\end{equation*}
Then, conditional on the data, the $Z_{nj}^{*2}$ are independent can
be expected to obey an \lln, so %
$\sum_j (Z_{nj}^{*2} - \E^* Z_{nj}^{*2}) \to^p 0$ %
and the CLT for the bootstrapped array requires $\sum_j \E^*
Z_{nj}^{*2}$ to converge to a positive and finite limit.  But, since
$\E^*$ only averages over the starting point of each block, we have
\begin{align*}
  \E^* Z_{nj}^{*2} &=
  \tfrac{1}{n} \sum_{\tau = 0}^{n-M_n}
  \Big(\tfrac{1}{\sqrt{n}}
  \sum_{t= \tau + 1}^{\tau + M_n} (X_{nt} - \bar X_n) \Big)^2 \\
  &= \tfrac{1}{n} \sum_{\tau_0 = 0}^{M_n-1}
  \sum_{j=1}^{J_n}
  \Big(\tfrac{1}{\sqrt{n}}
  \sum_{t= (j - 1) M_n + \tau_0 + 1}^{j M_n + \tau_0} (X_{nt} - \bar X_n) \Big)^2
\end{align*}
after grouping blocks separated by $M_n$ periods. For each $\tau_0$,
the summation can be expected to converge in probability through the
same arguments that were used to establish the \clt\ for the original
array.%
\footnote{Some of the details of the argument will typically need to
  change because the original \clt\ only requires convergence for
  $\tau_0 = 0$, but these details are often incidental to the original
  argument.}%
\footnote{Lemmas~\ref{L3} and
  \ref{L5} are particularly
  strong demonstrations of this argument.} %
A similar representation is available for the Circular and Stationary
bootstraps.

@@

\begin{thm}\label{T2}
Suppose that the conditions of Theorem~\ref{T1} hold
and let $\dist$ be any distance function that metricizes weak
convergence. Then
\begin{equation}\label{eq:7}
\pr^{*}[\dist(W_n^{*}, \sigma W) > \delta] \to^p 0
\end{equation}
for any positive  $\delta$.
If, in addition, $\sup_{t=1,\dots,n}|\mu_{nt} -
\bar \mu_n| = o(1/\sqrt{n})$ and
\begin{equation}\label{eq:4}
n^{-1} \sum_{s,t=1}^{\lfloor \gamma n \rfloor} \cov(X_{ns}, X_{nt}) \to \sigma^2 \gamma
\end{equation}
for all $\gamma \in [0,1]$, then
\begin{equation}\label{eq:8}
\pr[\dist(W_n, \sigma W) > \delta] \to 0
\end{equation}
for any positive $\delta$.
\end{thm}
If both~\eqref{eq:7} and~\eqref{eq:8} hold then the bootstrap can be
used to approximate the distribution of partial sums.
Note that Theorem~\ref{T2} imposes
stronger assumptions for the original process
than for the bootstrapped process. Without~\eqref{eq:4}, the
partial sum of the original observations converges to a transformed
Brownian Motion with a different covariance process. The bootstrapped
partial sum, on the other hand, always converges to standard (but
potentially scaled) Brownian Motion because the resampling strategies
ensure that the bootstrapped process is globally covariance stationary.

@@

\subsection*{Proof of Theorem~\ref{T1}}
{%
\newcommand{\Db}[1][j]{D_{n#1}^{*}}
\newcommand{\Zsum}{\sum_{j=1}^{J_n} \Zb / \sigma_n}
First we prove that
\begin{equation}\label{eq:37}
  \sup_x \big\lvert \pr^{*}\big[
  \sqrt{n}(\bar X_{n}^{*} -  \bar X_n) \big/ \sigma_n^{*}
  \leq x \big] - \Phi(x) \big\rvert \to^p 0
\end{equation}
where $\sigma_n^{*2} = n \E^* (\bar X_n^* - \bar X_n)^2$.
Rewrite $\sqrt{n}(\bar X_{n}^{*} - \bar X_n)$ as in
Equation~\eqref{eq:16}, so
\begin{equation*}
  \tfrac{1}{\sqrt{n}} \sum_{t=1}^n (X_{nt}^* - \bar X_n) =
  \sum_{j=1}^{J_n} Z_{nj}^*
\end{equation*}
and $\{\Zb / \sigma_n^*, \Gs_{nj}\}$ is a martingale difference
array. Moreover,
\begin{equation}
  \label{eq:18}
  \Pm^{*} \Big[\sum_{j=1}^{J_n} \Zb / \sigma_n^* \leq x \Big] - \Phi(x) \to^p 0,
\end{equation}
for all $x$ if $\sigma_n^{*2} \to^p \sigma^2$ (which ensures that
$\sigma_n^{*2}$ is uniformly a.s.\ positive and holds by
Lemma~\ref{L3}) and the following
two conditions hold for all positive $\epsilon$:
\begin{equation}
  \label{eq:19}
  \sum_{j=1}^{J_n} \Em^{*} \big(Z_{nj}^{*2} \ind\{Z_{nj}^{*2}  >
  \epsilon\}\big) \to^p 0
\end{equation}
and
\begin{equation}
  \label{eq:20}
  \Pm^{*}\Big[\ \Big|\sum_{j=1}^{J_n} Z_{nj}^{*2} - \sigma_n^{*2}
  \Big|\ > \epsilon \Big] \to^p 0
\end{equation}
since~\eqref{eq:19} and~\eqref{eq:20} ensure that $\Zb/\sigma_n^*$ obeys
a martingale difference \clt\ \citep[e.g.][Theorem 3.3]{HaH:80}.%
\footnote{Conditional on $X_{n1},\dots,X_{nn}$, $J_n$, and
  $M_{n1},\dots,M_{n,J_n}$, the only stochastic components of $\Zsum$
  are the start periods of each block, which are
  $\discreteuniform(1,\dots,n)$ and are independent of all of the
  other random variables in the information set used for conditioning.
  Consequently, $\Pm^*$ is a regular conditional probability and
  arguments like Hall and Heyde's (1980) Theorem 3.3
  apply without modification on this probability measure. See
  also Section 23.2 of \citet{Vaa:00}.
  Also note that Hall and Heyde's Theorem 3.3 as stated imposes an
  additional restriction on the sigma-fields. However, as Hall and
  Heyde discuss on pages 59 and 63--64, that condition is unnecessary
  here because $\sigma_n^{*2}$ is measurable with respect to
  all of the $\Gs_{nj}$.} %

For~\eqref{eq:20}, $Z_{nj}^*$ and $Z_{nk}^*$ (when $k \neq j$) are
conditionally uncorrelated given $X_{n1},\dots,X_{nn},$ and $\Ms$, which
implies
\begin{equation*}
  \sum_{j=1}^{J_n} Z_{nj}^{*2} - \sigma_n^{*2} =
  \sum_{j=1}^{J_n} \Big( Z_{nj}^{*2} - (1/J_n) \E^* \sum_{j=1}^{J_n} Z_{nj}^{*2} \Big)
\end{equation*}
almost surely. But
\begin{equation*}
   \Big\{\tfrac{n}{M_{nj}} \Big(Z_{nj}^{*2} - (1/J_n) \E^* \sum_{j=1}^{J_n} Z_{nj}^{*2}\Big),
   \ \Gs_{nj} \Big\}
\end{equation*}
is a uniformly-integrable martingale difference array
by Lemma~\ref{L5}
and satisfies the \lln\ (e.g., \citealp{Dav:94}, Theorem 19.7). So
this sum converges to zero in conditional probability, proving~\eqref{eq:20}.

>>>>

\newpage
\noindent Block bootstraps, e.g.\ the Moving Blocks
(\citealp{Kun:89}, and \citealp{LiS:92}), Circular Block \citep{PoR:92}, and Stationary
Bootstraps \citep{PoR:94}, have become popular in Economics, partly
because they do not require the researcher to make parametric
assumptions about the data generating process.  They are valid under
general weak dependence and moment conditions.  Some recent papers
(\citealp{GoW:02}, and \citealp{GoJ:03}) relax the dependence and moment
conditions of the original papers to fit with the Near-Epoch-Dependence
(\ned) assumptions commonly used in econometrics.%
\footnote{\citet{GoW:02} show that these bootstrap methods can be
  applied to heterogeneous $L_{2+\delta}$-\ned\ processes of size
  $-2(r-1)/(r-2)$ on a strong mixing sequence of size
  $-r(2+\delta)/(r-2)$, where $r > 2$ and $\delta >0$, when the
  original series has uniformly bounded $3r$-moments.  \citet{GoJ:03}
  relax these conditions to $L_{2+\delta}$-\ned\ of size $-1$ and
  $r+\delta$ moments for the original series, and size
  $-(2+\delta)(r+\delta)/(r-2)$ for the underlying mixing series.
  Both papers require that the expected block length grow with $n$ and
  be $o(n^{1/2})$.  \cite{GoP:11} discuss these issues further.}%
\footnote{An array $\{X_{nt}\}$ is an $L_{\rho}$-\ned\ process on a
  mixing array $\{V_{nt}\}$ if
  \begin{equation}
    \| X_{nt} - \E(X_{nt}
    \mid V_{n,t-m},\dots,V_{n,t+m}) \|_{\rho} \leq d_{nt} v_m
  \end{equation}
  with $v_m \to 0$ as $m \to \infty$ and $\{d_{nt}\}$ an array of
  positive constants.  It is of size $-\gamma$ if $v_m = O(m^{-\gamma
    - \delta})$ for all $\delta>0$.  Dropping the index ``$n$'' gives
  the series definition.  Note that the underlying strong and uniform mixing arrays
  are not required to be stationary.} %
But these conditions are still stronger than
required for a \clt\ to hold; \citet{Jon:97} has established the \clt\
under $L_2$-\ned\ with smaller size and moment
restrictions.%
\footnote{\citet{Jon:97} proves that the \clt\ holds for averages of
  $L_2$-\ned\ processes of size $-1/2$ on a strong mixing series of
  size $-r/(r-2)$, $r > 2$ and the original series having bounded
  $r$-moments.} %
This paper shows that these block bootstrap
methods consistently estimate the distribution of the sample mean
under \citepos{Jon:97} assumptions and show that an \fclt\ holds as
well.%
\footnote{\citet{Rad:96} proves consistency for the Moving Blocks
  Bootstrap for any stationary strong mixing sequence that satisfies
  the \clt. This paper uses a similar method of proof to his, but also
  accommodates nonstationary sequences and the Stationary
  Bootstrap.} %
It also relaxes \citepos{GoW:02} and \citepos{GoJ:03} requirement that
the expected block length be $o(n^{1/2})$ to the original papers'
requirement that it be $o(n)$, although their stronger assumption is
necessary for the bootstrap to be consistent under some forms of
heterogeneity.

@@

\begin{thm}\label{T1}
  Suppose the following conditions hold.
  \begin{enumerate}
  \item $X_{nt}$ is $L_2$-\ned\ of size $-1/2$ on an array
    $\{V_{nt}\}$ that is either strong mixing of size $-r/(r-2)$ or
    uniform mixing of size $-r/2(r-1)$, with $r > 2$.  The
    \ned\ magnitude indices are denoted $\{d_{nt}\}$.
  \item The array $\mu_{nt} - \bar \mu_n$ is uniformly bounded
    and $\sum_{t=1}^n (\mu_{nt} - \bar \mu_n)^2 = o(n p_n^2)$ (for
    the Stationary Bootstrap) or
    $\sum_{t=1}^n (\mu_{nt} - \bar \mu_n)^2 = o(n/M_n^2)$ (for
    the Moving or Circular Block Bootstraps),
    where $\E X_{nt} = \mu_{nt}$, $\bar{\mu}_n = n^{-1} \sum_{t=1}^n
    \mu_{nt}$, and $p_n$ and $M_n$ are further defined in item 4.
    Moreover, $\sqrt{n} \| \bar{X}_{n} - \bar\mu_n \|_2
    \to \sigma > 0$, with $\bar X_n = (1/n) \sum_{t=1}^n X_{nt}$.
  \item There exists an array of positive real numbers $\{c_{nt}\}$
    such that $(X_{nt} - \mu_{nt})/c_{nt}$ is uniformly $L_r$-bounded
    and $c_{nt}$ and $d_{nt}/c_{nt}$ are uniformly bounded in $n$
    and~$t$.
  \item $X_{nt}^{*}$ is generated by the Stationary Bootstrap with
    geometric block lengths with success probability $p_n$, $p_n = c
    n^{-a}$ and $a,c \in (0,1)$, or by the Moving or Circular Block
    bootstrap with block length $M_n$ such that $M_n \sim n^a$ for
    $a \in (0,1)$.  Let $M_{ni}$ be the block length of the $i$th
    block, $i=1,\dots,J_n$, and define $K_{n0} = 0$ and $K_{nj} =
    \sum_{i=1}^j M_{ni}$. The last block, $M_{n,J_n}$, is defined as
    $K_{n,J_n} - K_{n,J_n-1}$, so $K_{n,J_n} = n$ a.s.
  \end{enumerate}
  Then $\sigma^{*2} \to^p \sigma^2$, $\hat\sigma^{*2} \to^p \sigma^2$,
  \begin{equation}\label{eq:36}
    \sup_x \big\lvert \pr^{*}\big[\sqrt{n}(\bar X_{n}^{*} - \E^* \bar X_n^*) \leq x \big]
    - \pr \big[\sqrt{n}(\bar X_{n} - \E \bar X_n) \leq x \big] \big\rvert \to^p 0,
  \end{equation}
  and
  \begin{equation}
    \label{eq:2}
    \sup_x \big\lvert \pr^{*}\big[
    \sqrt{n}(\bar X_{n}^{*} -  \E^* \bar X_n^*) \big/ \hat\sigma_n^{*}
    \leq x \big] - \Phi(x) \big\rvert \to^p 0,
  \end{equation}
  where $\Phi$ is the \allcaps{cdf} of the Standard Normal
  distribution and
    \begin{equation}
      \label{eq:3}
      \hat{\sigma}_n^{*2} = \tfrac{1}{n} \sum_{j=1}^{J_n}
      \Big\{\sum_{t=K_{n,j-1}+1}^{K_{nj}} (X_{nt}^{*} - \bar X_{n}^{*})\Big\}^2.
    \end{equation}
\end{thm}

As mentioned earlier, these are the same size and mixing conditions
used by \citet{Jon:97}.
Note that De Jong does allow a little bit more flexibility in the
conditions on the array $\{c_{nt}\}$ \citep[see also][]{Dav:93};
essentially, he allows for a single set of blocks with the
maximal $\{c_{nt}\}$ over each block well-behaved, while this
paper requires this condition to hold for every possible partition of
blocks.  This additional restriction is required because the
Stationary Bootstrap will select the blocks randomly and is similar to
\citepos{JoD:00b} requirement for the \fclt.

The assumption on the dispersion of the individual means,
$(\mu_{nt} - \bar\mu_n)^2$, is the same as \citepos{GoW:02} and
\citepos{GoJ:03} when the expected block length is $o(\sqrt{n})$. When
the blocks are longer than $o(\sqrt{n})$, a case that \citet{GoW:02}
and \citet{GoJ:03} do not consider, the dispersion must be smaller.
To see these implications of this assumption, it is helpful to
consider a simple example: a stochastic process with a single break of
size $\Delta_n$ at time $T_n$,
\begin{equation*}
\mu_{nt} =
\begin{cases}
  \mu & \text{if } t \leq T_n \\
  \mu + \Delta_n & \text{otherwise}
\end{cases}
\end{equation*}
A small amount of algebra gives the relationship,
\begin{equation*}
  \sum_{t=1}^n (\mu_{nt} - \bar \mu_n)^2
  = T_n (n - T_n) \Delta_n^2 / n \equiv D_n;
\end{equation*}
we have defined $D_n$ to minimize notation in the subsequent
discussion.  For the Moving and Circular Block Bootstraps,
Theorem~\ref{T1}'s Assumption 2 holds if $D_n = o(n / M_n^2)$, and for
the Stationary Bootstrap the same relationship must hold for the
expected block length. If $\Delta_n = \Delta_0$ and $T_n = \pi n$ for
some constants $\Delta_0$ and $\pi$ and for all $n$, then
$D_n \approx \Delta_0^2 \, c (1 - c) \, n$ and Assumption~2 fails to
hold. These values of $\Delta_n$ and $T_n$ describe a scenario where
there is a break large enough that its timing could be precisely
estimated \citep[as in][]{Bai:94}, but the break is ignored and the
bootstrap is applied to the entire dataset.

We can also use the same DGP to produce scenarios where
Theorem~\ref{T1}'s Assumption~2 is satisfied.  The first mimics a
setting where the break is estimated and the bootstrap is applied to
the post-break sample, but a small number of pre-break observations
are accidentally included in the data set as well. Here
$\Delta_n = \Delta_0$ and $T_n = T_0$ for some finite $T_0$. In this
case, Assumption 2 holds hold as long as $M_n = o(\sqrt{n})$.
Similarly, if the break is moderate with unknown timing in the sense
of \citet{ElM:07}, so $\Delta_n = \Delta / \sqrt{n}$ and
$T_n = \pi n$, then $D_n \approx \Delta c (1 - c)$ and Assumption 2
again holds as long as $M_n = o(\sqrt{n})$. Consequently, block sizes
of length $o(\sqrt{n})$ ensure that the conditions of
Theorem~\ref{T1}'s when there is a moderate amount of
heterogeneity---either because the sample is chosen to lie after a
large break estimated break or because any breaks are smaller in
magnitude and, consequently, potentially difficult to detect. When the
degree of heterogeneity is very small or nonexistent, so
$\sum_t (\mu_{nt} - \bar \mu_n)^2$ grows even slower, then larger
block lengths can also satisfy Assumption 2.

@@

Our insight is that the expectation of squared blocks of the bootstrap
process can be expressed as a sequence of \emph{contiguous} blocks of
the original process, so the arguments that establish convergence of
the original squared blocks can be applied with only minor changes to
the bootstrapped blocks. Consider the Moving Blocks Bootstrap,%
\footnote{To make this presentation as simple as possible, assume for
  now that $n = M_n J_n$ exactly.} %
for example, and let
\begin{equation*}
  Z_{nj}^* = \tfrac{1}{\sqrt{n}} \sum_{t=(j-1) M_n + 1}^{j M_n} (X_{nt}^* - \bar X_n).
\end{equation*}
Then, conditional on the data, the $Z_{nj}^{*2}$ are independent can
be expected to obey an \lln, so %
$\sum_j (Z_{nj}^{*2} - \E^* Z_{nj}^{*2}) \to^p 0$ %
and the CLT for the bootstrapped array requires $\sum_j \E^*
Z_{nj}^{*2}$ to converge to a positive and finite limit.  But, since
$\E^*$ only averages over the starting point of each block, we have
\begin{align*}
  \E^* Z_{nj}^{*2} &=
  \tfrac{1}{n} \sum_{\tau = 0}^{n-M_n}
  \Big(\tfrac{1}{\sqrt{n}}
  \sum_{t= \tau + 1}^{\tau + M_n} (X_{nt} - \bar X_n) \Big)^2 \\
  &= \tfrac{1}{n} \sum_{\tau_0 = 0}^{M_n-1}
  \sum_{j=1}^{J_n}
  \Big(\tfrac{1}{\sqrt{n}}
  \sum_{t= (j - 1) M_n + \tau_0 + 1}^{j M_n + \tau_0} (X_{nt} - \bar X_n) \Big)^2
\end{align*}
after grouping blocks separated by $M_n$ periods. For each $\tau_0$,
the summation can be expected to converge in probability through the
same arguments that were used to establish the \clt\ for the original
array.%
\footnote{Some of the details of the argument will typically need to
  change because the original \clt\ only requires convergence for
  $\tau_0 = 0$, but these details are often incidental to the original
  argument.}%
\footnote{Lemmas~\ref{L4} and
  \ref{L6} are particularly
  strong demonstrations of this argument.} %
A similar representation is available for the Circular and Stationary
bootstraps.

@@

\begin{thm}\label{T2}
Suppose that the conditions of Theorem~\ref{T1} hold
and let $\dist$ be any distance function that metricizes weak
convergence. Then
\begin{equation}\label{eq:7}
\pr^{*}[\dist(W_n^{*}, \sigma W) > \delta] \to^p 0
\end{equation}
for any positive $\delta$. If, in addition,
\begin{equation}\label{eq:4}
n^{-1} \sum_{s,t=1}^{\lfloor \gamma n \rfloor} \cov(X_{ns}, X_{nt}) \to \sigma^2 \gamma
\end{equation}
for all $\gamma \in [0,1]$, then
\begin{equation}\label{eq:8}
\pr[\dist(W_n, \sigma W) > \delta] \to 0
\end{equation}
for any positive $\delta$.
\end{thm}
If both~\eqref{eq:7} and~\eqref{eq:8} hold then the bootstrap can be
used to approximate the distribution of partial sums.
Note that Theorem~\ref{T2} imposes
stronger assumptions for the original process
than for the bootstrapped process. Without~\eqref{eq:4}, the
partial sum of the original observations converges to a transformed
Brownian Motion with a different covariance process. The bootstrapped
partial sum, on the other hand, always converges to standard (but
potentially scaled) Brownian Motion because the resampling strategies
ensure that the bootstrapped process is globally covariance stationary.

@@

\subsection*{Proof of Theorem~\ref{T1}}
{%
\newcommand{\Db}[1][j]{D_{n#1}^{*}}
\newcommand{\Zsum}{\sum_{j=1}^{J_n} \Zb / \sigma_n}
Observe that
\[
\sqrt{n}(\bar X_{n}^{*} -  \E^{*} \bar X_n^{*})
= \sqrt{n}((\bar{X}_{n}^{*} - \bar{\mu}_{n}^{*}) - \E^{*} (\bar X_n^{*} - \bar{\mu}_n^{*}))
+ \sqrt{n}(\bar{\mu}_{n}^{*} - \E^{*} \bar{\mu}_n^{*}).
\]
Lemma~\ref{L3} implies that the last term is $o_{L_2}(1)$, so we can assume that
$\mu_{nt}$ is constant in $t$ for each $n$ for the proof of normality. Convergence
in probability of the bootstrap variance and the bootstrap variance estimator
holds by Lemma~\ref{L4}.

First we prove that
\begin{equation}\label{eq:37}
  \sup_x \big\lvert \pr^{*}\big[
  \sqrt{n}(\bar X_{n}^{*} -  \bar X_n) \big/ \sigma_n^{*}
  \leq x \big] - \Phi(x) \big\rvert \to^p 0
\end{equation}
where $\sigma_n^{*2} = n \E^* (\bar X_n^* - \bar X_n)^2$.
Rewrite $\sqrt{n}(\bar X_{n}^{*} - \bar X_n)$ as in
Equation~\eqref{eq:16}, so
\begin{equation*}
  \tfrac{1}{\sqrt{n}} \sum_{t=1}^n (X_{nt}^* - \bar X_n) =
  \sum_{j=1}^{J_n} Z_{nj}^*
\end{equation*}
and $\{\Zb / \sigma_n^*, \Gs_{nj}\}$ is a martingale difference
array. Moreover,
\begin{equation}
  \label{eq:18}
  \Pm^{*} \Big[\sum_{j=1}^{J_n} \Zb / \sigma_n^* \leq x \Big] - \Phi(x) \to^p 0,
\end{equation}
for all $x$ if $\sigma_n^{*2} \to^p \sigma^2$ (which ensures that
$\sigma_n^{*2}$ is uniformly a.s.\ positive and holds by
Lemma~\ref{L4}) and the following
two conditions hold for all positive $\epsilon$:
\begin{equation}
  \label{eq:19}
  \sum_{j=1}^{J_n} \Em^{*} \big(Z_{nj}^{*2} \ind\{Z_{nj}^{*2}  >
  \epsilon\}\big) \to^p 0
\end{equation}
and
\begin{equation}
  \label{eq:20}
  \Pm^{*}\Big[\ \Big|\sum_{j=1}^{J_n} Z_{nj}^{*2} - \sigma_n^{*2}
  \Big|\ > \epsilon \Big] \to^p 0
\end{equation}
since~\eqref{eq:19} and~\eqref{eq:20} ensure that $\Zb/\sigma_n^*$ obeys
a martingale difference \clt\ \citep[e.g.][Theorem 3.3]{HaH:80}.%
\footnote{Conditional on $X_{n1},\dots,X_{nn}$, $J_n$, and
  $M_{n1},\dots,M_{n,J_n}$, the only stochastic components of $\Zsum$
  are the start periods of each block, which are
  $\discreteuniform(1,\dots,n)$ and are independent of all of the
  other random variables in the information set used for conditioning.
  Consequently, $\Pm^*$ is a regular conditional probability and
  arguments like Hall and Heyde's (1980) Theorem 3.3
  apply without modification on this probability measure. See
  also Section 23.2 of \citet{Vaa:00}.
  Also note that Hall and Heyde's Theorem 3.3 as stated imposes an
  additional restriction on the sigma-fields. However, as Hall and
  Heyde discuss on pages 59 and 63--64, that condition is unnecessary
  here because $\sigma_n^{*2}$ is measurable with respect to
  all of the $\Gs_{nj}$.} %

For~\eqref{eq:20}, $Z_{nj}^*$ and $Z_{nk}^*$ (when $k \neq j$) are
conditionally uncorrelated given $X_{n1},\dots,X_{nn},$ and $\Ms$, which
implies
\begin{equation*}
  \sum_{j=1}^{J_n} Z_{nj}^{*2} - \sigma_n^{*2} =
  \sum_{j=1}^{J_n} \Big( Z_{nj}^{*2} - (1/J_n) \E^* \sum_{j=1}^{J_n} Z_{nj}^{*2} \Big)
\end{equation*}
almost surely. But
\begin{equation*}
   \Big\{\tfrac{n}{M_{nj}} \Big(Z_{nj}^{*2} - (1/J_n) \E^* \sum_{j=1}^{J_n} Z_{nj}^{*2}\Big),
   \ \Gs_{nj} \Big\}
\end{equation*}
is a uniformly-integrable martingale difference array
by Lemma~\ref{L6}
and satisfies the \lln\ (e.g., \citealp{Dav:94}, Theorem 19.7). So
this sum converges to zero in conditional probability, proving~\eqref{eq:20}.
